---
title: "Obesity_classification"
author: "Farah Salahuddin"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset Description and Exploratory Analysis

  The dataset obtained from UC Irvine Machine Learning Repository is called ‘Estimation of Obesity Levels Based on Eating Habits and Physical Condition’. The dataset consists of 2111 unique observations and 17 attributes that are as follows:
Gender, Age, Height, Weight, Family History, if the person eats high caloric food frequently, if the person eats vegetables in meals, number of meals in a day, if the individual eats between meals, smoking, amount of water intake, frequency of physical activity, time on technological devices, frequency of alcohol intake and mode of transportation. Out of these, some are numeric variables for example age, weight and height, while others are categorical variables such as gender, smoking and transportation. 

The relationship between the predictor variables and the response variable, obesity level was studied.

```{r}

library(ggcorrplot)
library(ggplot2)
library(gridExtra)
library(pls)
library(glmnet)
library(MASS)
library(leaps)
library(dplyr)
library(tidyr)
library(readxl)
library(rpart)
library(randomForest)
library(caret)
library(gbm)
library(adabag)
library(patchwork)
```




```{r}
url <- "https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip"
temp_zip <- tempfile(fileext = ".zip")
download.file(url, temp_zip, mode = "wb")


```





```{r}
unzip_dir <- tempdir()
unzip(temp_zip, exdir = unzip_dir)


```


```{r}
list.files(unzip_dir)
file_path <- file.path(unzip_dir, "ObesityDataSet_raw_and_data_sinthetic.csv")
obesity <- read.table(file_path,sep=",",header=TRUE)


```

```{r}
head(obesity)

```


```{r}
obesity$Gender=factor(obesity$Gender)
obesity$family_history_with_overweight=factor(obesity$family_history_with_overweight)
obesity$FAVC=factor(obesity$FAVC) #high caloric food
obesity$FCVC=factor(obesity$FCVC) #vegetables
obesity$FCVC=round(as.numeric(levels(obesity$FCVC)[obesity$FCVC]))
obesity$FCVC=factor(obesity$FCVC) #vegetables

obesity$FAF=factor(obesity$FAF) #physical activity
obesity$FAF=round(as.numeric(levels(obesity$FAF)[obesity$FAF]))
obesity$FAF=factor(obesity$FAF) #physical activity


obesity$CAEC=factor(obesity$CAEC) #eat any food between meals
obesity$SMOKE=factor(obesity$SMOKE)
obesity$SCC=factor(obesity$SCC) #monitor calories you eat
obesity$CALC=factor(obesity$CALC) #alcohol
obesity$MTRANS=factor(obesity$MTRANS) #transportation
obesity$NObeyesdad=factor(obesity$NObeyesdad)

```


```{r}

names_ob=c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")
ggplot(obesity,aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")) 
                   ,fill=NObeyesdad
                    )) + geom_bar(fill="lightblue") + 
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Nobeyesdad',title = "Counts of different levels of obesity in data", x = "Obesity Category", y = "Frequency")

```


```{r}
colnames_predictors=c("Gender","family_history_with_overweight","FAVC","FCVC","CAEC","SMOKE","SCC","CALC","MTRANS")
colnames_pred_num=c("Age","Height","Weight","NCP")

data_long <- obesity %>%
  pivot_longer(cols = colnames_pred_num, names_to = "Variable", values_to = "Value")

head(data_long)

ggplot(obesity, aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")), y = Age, fill = NObeyesdad)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Obesity Level',title = "Distribution of Obesity Levels by Age", x = "Obesity Category", y = "Age")



ggplot(obesity, aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")), y = Weight, fill = NObeyesdad)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Obesity Level',title = "Distribution of Weight by Obesity Category", x = "Obesity Category", y = "Weight")


ggplot(obesity, aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")), y = Height, fill = NObeyesdad)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Obesity Level', title = "Distribution of Height by Obesity Category", x = "Obesity Category", y = "Height")

ggplot(obesity, aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")), y = NCP, fill = NObeyesdad)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Obesity Level',title = "Relationship between Number of Meals Eaten Per Day and Obesity", x = "Obesity Category", y = "Number of meals eating per day") + 
  coord_flip()


ggplot(obesity, aes(x = MTRANS, fill =NObeyesdad)) +
  geom_bar(position = "fill") +  # "fill" makes proportions (100% stacked)
  labs(fill='Obesity Level',
       title = "Relationship between Mode of transportation frequently used and obesity level", 
       x = "Mode of Transportation", 
       y = "Proportion") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```



```{r}

colnames_predictors=c("Gender","family_history_with_overweight","FAVC","FCVC","CAEC","SMOKE","SCC","CALC","MTRANS")


p1<-ggplot(obesity, aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")), y = Age, fill = NObeyesdad)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Obesity Level', x = "Obesity Category", y = "Age")



p2<-ggplot(obesity, aes(x = factor(NObeyesdad, levels = c("Insufficient_Weight", "Normal_Weight", "Overweight_Level_I", "Overweight_Level_II","Obesity_Type_I","Obesity_Type_II","Obesity_Type_III")), y = Height, fill = NObeyesdad)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="none",axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(fill='Obesity Level', x = "Obesity Category", y = "Height")


p3<-ggplot(obesity, aes(x = MTRANS, fill =NObeyesdad)) +
  geom_bar(position = "fill") +  # "fill" makes proportions (100% stacked)
  labs(fill='Obesity Level',
       #title = "Relationship between Mode of transportation frequently used and obesity level", 
       x = "Mode of Transportation", 
       y = "Proportion") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p4<-ggplot(obesity, aes_string(x=obesity$FAF,fill=obesity$NObeyesdad)) +
    geom_bar(position="fill") +
    scale_fill_brewer(palette="Set2") +
    labs(fill="Obesity Level",
         #title="Relationship between physical activity and obesity",
         x="Frequency of Physical Activity", y="Obesity Levels") + 
    scale_x_discrete(labels=c("0"="Never","1"="1 to 2 days","2"="2 to 4 days","3"="4 to 5 days")) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  

(p1 | p2) / (p3 | p4)

#do.call(grid.arrange, c(plots, ncol = 2)) 

#title("Relationship between obesity level and different categorical feature variables")


```

```{r}

ggplot(obesity, aes_string(x=obesity$FAF,fill=obesity$NObeyesdad)) +
    geom_bar(position="fill") +
    scale_fill_brewer(palette="Set2") +
    labs(fill="Obesity Level",title="Relationship between physical activity and obesity",x="Frequency of Physical Activity", y="Obesity Levels") + 
    scale_x_discrete(labels=c("0"="Never","1"="1 to 2 days","2"="2 to 4 days","3"="4 to 5 days")) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  

```